Supplemental Materials for ``Universals of word order result from optimization of grammars for efficient communication''

### 2019-09-24 writeup/pnas-word-order-si.tex

@@

\paragraph{Inferring $\lambda$}
We choose $\lambda$ to minimize the number of baselines that have better values of $R_\lambda := R_{Pars} + \lambda R_{Pred}$ than the real language:
\begin{equation}
	O := \min_\lambda \# \{b \in B: R_{eff}(b) > R_{eff}(r)\}
\end{equation}
where $r$ is the real language, and $B$ is the set of baselines.
Naively choosing $\lambda$ to minimize this quantity from the observed $50$ baseline samples will bias $O$ towards overestimating the optimality of the real language.

TODO maybe add a figure explaining the Bayesian story a bit more -- how we're using a continuous distribution to smooth out the observed samples?

We eliminate this anti-conservative bias by conducting joint Bayesian inference of $\lambda$ of $O$, using the Dirichlet Gaussian Process, a Bayesian version of Kernel Density Smoothing.
The 50 samples represent an underlying baseline distribution in the predictability-parseability plane.
We approximate this underlying distribution as a continuous distribution from the Dirichlet Gaussian Process, and infer its parameters using Bayesian inference.
We use this method to obtain a Bayesian estimate of the distribution of baseline distributions in the Surprisal-Parseability plane.
The Dirichlet Gaussian Process (CITE) defines a prior over possible baseline distributions.
Under this prior, each possible distribution is parameterized by a number $n$ of components, and $n$ Gaussians $\mu_n, \sigma^2_n$.
Due to the flexible number of components, this nonparametric prior can accomodate a wide range of continuous probability distributions, including multimodal ones.
We use the $50$ observed baseline samples to infer a posterior distribution over such smooth distributions.
We then use MCMC to obtain 1,000 samples from the posterior, based on these samples.
We used the MCMC implementation provided by the \texttt{R} \citep{R2013} package \texttt{dirichletprocess}~\citep{ross2018dirichletprocess}, discarding the first 1,000 samples as warmup samples, and obtaining another 1,000 samples.

For every sampled distribution -- consisting of a component number $n$ and $n$ Gaussians, we then computed 
\begin{equation}
	O := \min_\lambda P(R_{eff}(b) > R_{eff}(r))\}
\end{equation}
where we considered $\lambda \in (0,1)$ in steps of $0.5$.

By taking the values $O, \lambda$ for the 1,000 posterior samples, we obtained an estimate of the posterior of $O$ and $\lambda$.

>>>>

Two ways of quatifying optimality

@@

%\paragraph{Inferring $\lambda$}
%We choose $\lambda$ to minimize the number of baselines that have better values of $R_\lambda := R_{Pars} + \lambda R_{Pred}$ than the real language:
%\begin{equation}
%	O := \min_\lambda \# \{b \in B: R_{eff}(b) > R_{eff}(r)\}
%\end{equation}
%where $r$ is the real language, and $B$ is the set of baselines.
%Naively choosing $\lambda$ to minimize this quantity from the observed $50$ baseline samples will bias $O$ towards overestimating the optimality of the real language.
%
%TODO maybe add a figure explaining the Bayesian story a bit more -- how we're using a continuous distribution to smooth out the observed samples?
%
%We eliminate this anti-conservative bias by conducting joint Bayesian inference of $\lambda$ of $O$, using the Dirichlet Gaussian Process, a Bayesian version of Kernel Density Smoothing.
%The 50 samples represent an underlying baseline distribution in the predictability-parseability plane.
%We approximate this underlying distribution as a continuous distribution from the Dirichlet Gaussian Process, and infer its parameters using Bayesian inference.
%We use this method to obtain a Bayesian estimate of the distribution of baseline distributions in the Surprisal-Parseability plane.
%The Dirichlet Gaussian Process (CITE) defines a prior over possible baseline distributions.
%Under this prior, each possible distribution is parameterized by a number $n$ of components, and $n$ Gaussians $\mu_n, \sigma^2_n$.
%Due to the flexible number of components, this nonparametric prior can accomodate a wide range of continuous probability distributions, including multimodal ones.
%We use the $50$ observed baseline samples to infer a posterior distribution over such smooth distributions.
%We then use MCMC to obtain 1,000 samples from the posterior, based on these samples.
%We used the MCMC implementation provided by the \texttt{R} \citep{R2013} package \texttt{dirichletprocess}~\citep{ross2018dirichletprocess}, discarding the first 1,000 samples as warmup samples, and obtaining another 1,000 samples.
%
%For every sampled distribution -- consisting of a component number $n$ and $n$ Gaussians, we then computed 
%\begin{equation}
%	O := \min_\lambda P(R_{eff}(b) > R_{eff}(r))\}
%\end{equation}
%where we considered $\lambda \in (0,1)$ in steps of $0.5$.
%
%By taking the values $O, \lambda$ for the 1,000 posterior samples, we obtained an estimate of the posterior of $O$ and $\lambda$.
